\ttitle

### 2014-01-20 Chapters/Chapter3/one_class_svm.tex
Final 1.0

\begin{figure}
\centering
  \includegraphics[width=0.8\textwidth]{./Figures/chapter3/svm_mapping_spaces.pdf}
  \caption[Mapping spaces in \gls{svm}]{Schematic visualization of the non-linear mapping of \glspl{svm}.
  The input vector $\vectorsym{x}$ from input space $\mathcal{I}$ is mapped by a non-linear function $\phi(\vectorsym{x})$ to a feature vector $\vectorsym{z}$ in the high dimensional feature space $\mathcal{F}$.
  The weights of $\vectorsym{w}$ create a linear separating hyperplane, which maps the high dimensional vector to the predicted outcome $\vectorsym{y}$.}
  \label{fig:svm_mapping_spaces}
\end{figure}

@@

\begin{figure}
\centering
  \includegraphics[width=0.5\textwidth]{./Figures/chapter3/svm_separating_plane_with_margin.png}
  \caption[\gls{svm} and the separating hyperplane]{Illustration of the separating hyperplane of an \gls{svm}.
  Here $w$ is the normal vector for the separating hyperplane and size of the margin is $\frac{2}{\lVert{w}\rVert}$.
  Image from Wikipedia.org}
  \label{fig:svm_hyperplane}
\end{figure}

@@

\subsection{\acrlong{nu-svm}}\label{subsec:nu-svm}
The first of the \gls{oc-svm} methods we will discuss is often referred to as \gls{nu-svm} and introduced by Sch\"olkopf \etal \cite{scholkopf1999support}.
Instead of estimating the density function of an distribution $P$, it focuses on an easier problem: the algorithm finds regions in the input where the ``probability density lives''.
This results in a function such that most of the data is in the region where the function is non-zero.

@@

\begin{figure}
  \centering
    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/nu-svm.pdf}
  \caption[\gls{nu-svm}]{Graphical representation of \gls{nu-svm}. The separating hyperplane $w \cdot \phi(x_i) + \rho = 0$ create a maximal margin, in the feature space, between the data points and the origin. Slack variables $\xi_i$ are used to create a soft margin.}
  \label{fig:nu-svm}
\end{figure}

@@

\subsection{\acrlong{svdd}}\label{subsec:oc-svm-svdd}
The method introduced by Tax and Duin \cite{tax1999support}, known as \acrlong{svdd}, follows a spherical instead of planar approach.
The boundary, created in feature space $\mathcal{F}$, forms a hypersphere around the (high density region of the) data.
The volume of this hypersphere is minimized to get the smallest enclosing boundary.
The chance of accepting outlier objects is thereby also minimized \cite{tax2003online}.
By allowing outliers using slacks variables, in the same manner as classical \gls{svm} and \gls{nu-svm}, a soft margin is constructed.

\begin{figure}
  \centering
    \includegraphics[width=0.8\textwidth,keepaspectratio]{./Figures/chapter3/nu-vs-svdd.pdf}
  \caption[Difference \gls{nu-svm} and \gls{svdd}]{Graphical representation of the difference between \gls{nu-svm} (left) and \gls{svdd} (right). Note that for the sake of simplicity the kernel functions are not applied.}
  \label{fig:nu-vs-svdd}
\end{figure}

@@

\begin{figure}
  \centering
    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/svdd-boundary.pdf}
  \caption[\gls{svdd} boundary]{The \gls{svdd} method trained on a banana-shaped data set with different sigma-values for the \gls{rbf} kernel. Solid circles are support vectors, the dashed line is the boundary. Image by Tax \cite{tax2001one}.}
  \label{fig:svdd-boundary}
\end{figure}

@@

\begin{figure}
  \centering
    \includegraphics[width=0.8\textwidth,keepaspectratio]{./Figures/chapter3/svdd-parameter-sigma.pdf}
  \caption[\gls{svdd} boundary]{The \gls{svdd} method trained on a banana-shaped data set with different $\sigma$-values for the \gls{rbf} kernel. Solid circles are support vectors, the dashed line is the boundary.}
  \label{fig:svdd-boundary-sigma}
\end{figure}

>>>>

\begin{figure}
\centering
  \includegraphics[width=0.8\textwidth]{./Figures/chapter3/svm_mapping_spaces.pdf}
  \caption[Mapping spaces in Support Vector Machines]{Schematic visualization of the non-linear mapping of \glspl{svm}.
  The input vector $\vectorsym{x}$ from input space $\mathcal{I}$ is mapped by a non-linear function $\phi(\vectorsym{x})$ to a feature vector $\vectorsym{z}$ in the high dimensional feature space $\mathcal{F}$.
  The weights of $\vectorsym{w}$ create a linear separating hyperplane, which maps the high dimensional vector to the predicted outcome $\vectorsym{y}$.}
  \label{fig:svm_mapping_spaces}
\end{figure}

@@

\begin{figure}
\centering
  \includegraphics[width=0.5\textwidth]{./Figures/chapter3/svm_separating_plane_with_margin.png}
  \caption[Support Vector Machines and the separating hyperplane]{Illustration of the separating hyperplane of an \gls{svm}.
  Here $w$ is the normal vector for the separating hyperplane and size of the margin is $\frac{2}{\lVert{w}\rVert}$.
  Image from Wikipedia.org}
  \label{fig:svm_hyperplane}
\end{figure}

@@

\subsection{\texorpdfstring{$\nu$-Support Vector Machine}{nu-Support Vector Machine}}\label{subsec:nu-svm}
The first of the \gls{oc-svm} methods we will discuss is often referred to as \gls{nu-svm} and introduced by Sch\"olkopf \etal \cite{scholkopf1999support}.
Instead of estimating the density function of an distribution $P$, it focuses on an easier problem: the algorithm finds regions in the input where the ``probability density lives''.
This results in a function such that most of the data is in the region where the function is non-zero.

@@

\begin{figure}
  \centering
    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/nu-svm.pdf}
  \caption[$\nu$-Support Vector Machine illustration]{Graphical representation of \gls{nu-svm}. The separating hyperplane $w \cdot \phi(x_i) + \rho = 0$ create a maximal margin, in the feature space, between the data points and the origin. Slack variables $\xi_i$ are used to create a soft margin.}
  \label{fig:nu-svm}
\end{figure}

@@

\subsection{\texorpdfstring{\acrlong{svdd}}{SVDD}}\label{subsec:oc-svm-svdd}
The method introduced by Tax and Duin \cite{tax1999support}, known as \acrlong{svdd}, follows a spherical instead of planar approach.
The boundary, created in feature space $\mathcal{F}$, forms a hypersphere around the (high density region of the) data.
The volume of this hypersphere is minimized to get the smallest enclosing boundary.
The chance of accepting outlier objects is thereby also minimized \cite{tax2003online}.
By allowing outliers using slacks variables, in the same manner as classical \gls{svm} and \gls{nu-svm}, a soft margin is constructed.

\begin{figure}
  \centering
    \includegraphics[width=0.8\textwidth,keepaspectratio]{./Figures/chapter3/nu-vs-svdd.pdf}
  \caption[Relation between $\nu$-SVM and SVDD]{Graphical representation of the difference between \gls{nu-svm} (left) and \gls{svdd} (right). Note that for the sake of simplicity the kernel functions are not applied.}
  \label{fig:nu-vs-svdd}
\end{figure}

@@

\begin{figure}
  \centering
    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/svdd-boundary.pdf}
  \caption[SVDD boundary with sigma parameter]{The \gls{svdd} method trained on a banana-shaped data set with different sigma-values for the \gls{rbf} kernel. Solid circles are support vectors, the dashed line is the boundary. Image by Tax \cite{tax2001one}.}
  \label{fig:svdd-boundary}
\end{figure}

@@

\begin{figure}
  \centering
    \includegraphics[width=0.8\textwidth,keepaspectratio]{./Figures/chapter3/svdd-parameter-sigma.pdf}
  \caption[SVDD boundary with sigma parameter]{The \gls{svdd} method trained on a banana-shaped data set with different $\sigma$-values for the \gls{rbf} kernel. Solid circles are support vectors, the dashed line is the boundary.}
  \label{fig:svdd-boundary-sigma}
\end{figure}

### 2014-01-20 Chapters/Chapter6/data_gathering.tex
Final 1.0

% ===== OUTDOOR FREE ACTIVITIES ========
\subsection{Outdoor free}\label{subsec:outdoor_free}
The second series of outdoor recordings consist of more free form activities.
The performed activities consist, as with the first set, of standing, walking, and running.
Furthermore, the segments are performed in a circular manner.
\eg the running activities are performed around a fountain with an estimated diameter of 10 meters.
The plots of the recordings of Subject 2, for the accelerometer, rotation, and magnetometer are shown in \Cref{fig:plots_subject_2}.
For Subject~$2$ the still of the video recordings are in the upper row of \Cref{fig:stills_subject_1_and_2}, for Subject~$1$ in the lower row.

@@

%---- Run 3: Stairs remco
\begin{figure}
  \centering
  \begin{subfigure}[b]{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_rot_annotated.eps}
    \caption{Run 7: Indoor stairs subject 3, rotation}
    \label{fig:data_gathering_run_7_rot}
  \end{subfigure} \\

  \begin{subfigure}[b]{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_mag_annotated.eps}
    \caption{Run 7: Indoor stairs subject 3, Mag}
    \label{fig:data_gathering_run_7_mag}
  \end{subfigure} \\

  \begin{subfigure}[b]{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_annotated.eps}
    \caption{Run 7: Indoor stairs subject 3, accelerometer}
    \label{fig:data_gathering_run_7_acc}
  \end{subfigure}
  \caption[Plots subject 3]{Annotated plots of a full recording performed by Subject~$3$ while walking the stairs. The vertical lines are manually determined change points. The label above each segment indicates the current performed activity.}\label{fig:plots_subject_3}
\end{figure}

>>>>

% ===== OUTDOOR FREE ACTIVITIES ========
\subsection{Outdoor free}\label{subsec:outdoor_free}
The second series of outdoor recordings consist of more free form activities.
The performed activities consist, as with the first set, of standing, walking, and running.
Furthermore, some segments are performed in a circular manner.
For example, the running activities are performed around a fountain with an estimated diameter of 10 meters.
The plots of the recordings of Subject 2, for the accelerometer, rotation, and magnetometer are shown in \Cref{fig:plots_subject_2}.
For Subject~$2$ the still of the video recordings are in the upper row of \Cref{fig:stills_subject_1_and_2}, for Subject~$1$ in the lower row.

@@

%---- Run 3: Stairs remco
\begin{figure}
  \centering
  \begin{subfigure}[b]{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_annotated.eps}
    \caption{Accelerometer}
    \label{fig:data_gathering_run_7_acc}
  \end{subfigure} \\

  \begin{subfigure}[b]{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_rot_annotated.eps}
    \caption{Rotation}
    \label{fig:data_gathering_run_7_rot}
  \end{subfigure} \\

  \begin{subfigure}[b]{1\textwidth}
    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_mag_annotated.eps}
    \caption{Magnetic field}
    \label{fig:data_gathering_run_7_mag}
  \end{subfigure}

  \caption[Plots subject 3]{Annotated plots of a full recording performed by Subject~$3$ while walking the stairs, indoor. The vertical lines are manually determined change points. The label above each segment indicates the current performed activity.}\label{fig:plots_subject_3}
\end{figure}

### 2014-01-20 Chapters/Chapter6/data_sets.tex
Final 1.0

Using the video recordings, the data sets are manually annotated, as shown in the plots of \Cref{fig:plots_subject_1,fig:plots_subject_2,fig:plots_subject_3}.
The graphs display each from a single recording the inertial sensor values (see \Cref{tab:recorded_metrics} for explanation of the metrics), with manually annotated activities above the data.
The manually determined change points are used to draw quality conclusion over the algorithm's performance, by comparing it to the discovered change points.
All of the used data sets, including the video material and manual labeling, are made available to the public as the \acrlong{almende-data} \cite{vlasveld2014acras} for further research.
The following section discusses the performed activities during the recordings.

>>>>

Using the video recordings, the data sets are manually annotated, as shown in the plots of \Cref{fig:plots_subject_1,fig:plots_subject_2,fig:plots_subject_3}.
The graphs display each from a single recording the inertial sensor values (see \Cref{tab:recorded_metrics} for an explanation of the metrics), with manually annotated activities above the data.
The manually and discovered change points are compared to draw quality conclusions over the algorithm's performance.
All of the used data sets, including the video material and manual labeling, are made available to the public as the \glsfirst{almende-data} \cite{vlasveld2014acras} for further research.
The following section discusses the performed activities during the recordings.

### 2014-01-20 Chapters/Chapter6/results.tex
Final 1.0

\subsection{Quantitative metrics}
The employed two-stage algorithm requires the setting of user defined parameters.
In \Cref{tab:results_final_real_world} for each run the results are listed.
The right-most column summarizes the overall performance by taking the average over all the runs.
We have used a window length of 50 frames, which represents around $1.16$ seconds of data (the frame-rate of the recordings is not fixed).
The other parameter values, \ie the \gls{rbf} $\sigma$, high and low threshold, and proximity parameter $\delta$, are the result of an educated guess and manual optimization.
In general, a high value for the upper threshold and low value for the lower threshold indicates a robust method, since it indicates that change points are significantly different from homogeneous segments.
The proximity $\delta$ should be low, since that indicates there are few false positives.

@@

\begin{table}
  \centering
  \caption[Results real world runs]{Results of the real-world data sets. The used parameters are: window length = $50$, $\sigma = 13$, $th_{high} = 1.2$, $th_{low} = 0.8$, $\delta = 0.7$.
  The table shows the \gls{far} $\gamma$, average delay $d_{avg}$ and its spread, using the standard deviation.
  Low values are better.
  }
  \begin{tabulary}{\textwidth}{|l|c|c|c|c|c|c|c||c|}
    \cline{2-9}
    \multicolumn{1}{l|}{} & Run 1 & Run 2 & Run 3 & Run 4 & Run 5 & Run 6 & Run 7 & Average \\
    \hline
    $\gamma(Y)$ & 0.1 & 0.2 & 0.14 & 0.37 & 0.29 & 0.64 & 0.57 & 0.33 \\
    \hline
    $d_{avg}$ & 0.52 & 0.85 & 0.93 & 0.54 & 0.69 & 0.51 & 0.32 & 0.62\\
    \hline
    $std(d_{avg})$ & 0.71 & 0.31 & 0.68 & 0.37 & 0.50 & 0.44 & 0.35 & 0.48 \\
    \hline
  \end{tabulary}
  \label{tab:results_final_real_world}
\end{table}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/boxplot_final_runs.eps}
  \caption[Box plot results real-world runs]{Box plot of the results for the real-world runs, indicating the number of data points between the actual and closest detected change points. A lower and more compact box plot is better.}
  \label{fig:boxplot_final_real_world_runs}
\end{figure}

@@

Overall our impression is that the proposed method is fairly good at recognizing change points.
Depending on the parameters for the sensitivity adjustment, almost all changes are discovered.
With a high sensitivity it seems, initially, the method produces many false positives.
After closer inspecting of the data, we can confirm that there is a change in the data, although the performed activity stayed the same.
This is also a weakness of the method: since the sensitivity is set globally, it can produce many semi-false positive change points for some segments of a run.

An other positive observation is that the method often recognizes a change point \emph{before} our annotation.
Again, after closer inspection we can see that the data changes shape and distribution before we annotated a change based on the video recordings.
This effect can be observed in the case of a transition from running to walking; the last `steps' of running are already recognized of being different.
As with the above behavior, this is subject to parameter settings.

Considering the overall performance and results, we have the following observations of cases in which the method has problems with detecting change:
\begin{itemize}
  \item \textbf{Merging:} in our method we use a proximity time period $\delta$ (usually up to $0.5$ - $1.0$ seconds) to merge discovered change points that have a very small distance between them.
  Multiple merging-strategies can be applied, and in our method we use the naive implementation by simply ignoring all the discovered change points that occur less than $\delta$ seconds after the previous change point.
  This works well when there is a block of noisy data with a high amount of (falsely) discovered change points.
  On the flip side, this method will ignore real new change points which occur during the noisy period.
  This problem on itself can be formulated as a change detection problem.
  \item \textbf{Masking:} Since the parameters for the detection method are set globally, it is difficult to discover all the change points (and only the change points) without a high \gls{far}.
  If a change point is proceeded by a noisy block, then the real change point will be merged with the previous change points.
  In an other case, where a change point is very clear represented in the data but the next, close by, other change point is more subtle, the latter change point is \emph{masked} by the former.
  This masking effect is also discussed in \cite{inclan1994use}, where an iterated approach is applied.
  \item \textbf{False heterogeneity:} For our video-data synchronization, we started and ended each recording with a few seconds in which the recording smartphone was kept still in the air.
  During this period, the data variance, and thus the constructed hypersphere, becomes very small.
  As a results, even small movements are considered to be changes and in the case of the mid-air still smartphone a lot of false change points are detected.
  Due to the merging effect described above, the final real change point (often shaking the smartphone) is not discovered.
  \item \textbf{Incorrect weighting:} In our analyses we have used the data from the accelerometer, magnetic field, and rotation sensors.
  In the segments which embodied movement in a circular manner (such as walking and running around a fountain), the turn was not (or at a different time point) discovered.
  Alternatively, when we only used the magnetic field and rotation sensors, the turn was correctly discovered.
  Other transitions, such as from walking to running, are harder to discover without the (linear) accelerometer sensor data.
\end{itemize}

An more elaborate discussion with observations and remarks is provided in \Cref{AppendixA}.

\section{Possible improvements}\label{sec:possible_improvements}
From the above section we can extract a few possible improvements for the algorithm.
Most of the observations and remarks are due to a different required sensitivity for local segments of the run.
Currently, our method employs a global setting for the thresholds.
A possible improvement would be to use locally optimized parameters, although that would require an reflective data processing approach or more a priori information about the data.
With our approach, the lack of a priori knowledge about the data distribution is one of the design decisions.

Furthermore, we discovered that sometimes, especially with movements incorporating circular motions, the addition of (linear) accelerometer data makes the discovery of change points harder.
In such cases, each of the inertial sensor data streams should be relatively weighted (\eg using the covariance) in order to signal an overall change when it only occurs in one of the streams.

>>>>

\subsection{Quantitative metrics}
The employed two-stage algorithm requires the setting of user defined parameters.
In \Cref{tab:results_final_real_world} for each run the results are listed.
The right-most column summarizes the overall performance by taking the average over all the runs.
We have used a window length of 50 frames, which represents around $1.16$ seconds of data (the frame-rate of the recordings is not fixed).
The other parameter values, \ie the \gls{rbf} $\sigma$, high and low threshold, and proximity parameter $\delta$, are the result of an educated guess and manual optimization.
In general, a high value for the upper threshold and low value for the lower threshold indicates a robust method, since it indicates that change points are significantly different from homogeneous segments.
The chosen proximity $\delta$ should be low, since that indicates there will be few false positives.

@@

\begin{table}[H]
  \centering
  \caption[Results real world runs]{Results of the real-world data sets. The used parameters are: window length = $50$, $\sigma = 13$, $th_{high} = 1.2$, $th_{low} = 0.8$, $\delta = 0.7$.
  The table shows the \gls{far} $\gamma$, average delay $d_{avg}$ and its spread, using the standard deviation.
  Low values are better.
  }
  \begin{tabulary}{\textwidth}{|l|c|c|c|c|c|c|c||c|}
    \cline{2-9}
    \multicolumn{1}{l|}{} & Run 1 & Run 2 & Run 3 & Run 4 & Run 5 & Run 6 & Run 7 & Average \\
    \hline
    $\gamma(Y)$ & 0.1 & 0.2 & 0.14 & 0.37 & 0.29 & 0.64 & 0.57 & 0.33 \\
    \hline
    $d_{avg}$ & 0.52 & 0.85 & 0.93 & 0.54 & 0.69 & 0.51 & 0.32 & 0.62\\
    \hline
    $std(d_{avg})$ & 0.71 & 0.31 & 0.68 & 0.37 & 0.50 & 0.44 & 0.35 & 0.48 \\
    \hline
  \end{tabulary}
  \label{tab:results_final_real_world}
\end{table}

\begin{figure}[H]
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/boxplot_final_runs.eps}
  \caption[Box plot results real-world runs]{Box plot of the results for the real-world runs, indicating the number of data points between the actual and closest detected change points. A lower and more compact box plot is better.}
  \label{fig:boxplot_final_real_world_runs}
\end{figure}

@@

Overall our impression is that the proposed method is fairly good at recognizing change points.
Depending on the parameters for the sensitivity adjustment, almost all changes are discovered.

Our first observation is that with a high sensitivity it seems that the method produces many false positives.
Although the performed activity did not change around those false positives, there was a change in the recorded data.
This is a drawback of the method: since the sensitivity is set globally, it can produce false positive change points for some segments of a run.

The second observation is that the method often recognizes a change point \emph{before} our annotation.
Again, after closer inspection we can see that the data changes shape and distribution before we annotated a change based on the video recordings.
This effect can be observed in the case of a transition from running to walking; the last `steps' of running are already labeled as change points.
As with the above behavior, this is subject to parameter settings.

Considering the overall performance and results, we have the following observations:
\begin{itemize}
  \item \textbf{Merging:} in our method we use a proximity time period $\delta$ (usually up to $0.5$ - $1.0$ seconds) to merge discovered change points that are close together.
  Multiple merging-strategies can be applied, and in our method we use the naive implementation by simply ignoring all the discovered change points that occur less than $\delta$ seconds after the previous change point.
  This works well when there is a block of noisy data with a high amount of (falsely) discovered change points.
  On the flip side, this method will ignore real new change points which occur during the noisy period.
  This problem on itself can be formulated as a change detection problem.
  \item \textbf{Masking:} Since the parameters for the detection method are set globally, it is difficult to discover all the change points (and only the change points) without a high false alarm rate.
  If a change point is proceeded by a noisy block, then the real change point will be merged with the previous change points.
  In an other case, where a change point is very clear represented in the data but the next near change point is more subtle, the latter change point is \emph{masked} by the former.
  This masking effect is also discussed in \cite{inclan1994use}, where an iterated approach is applied.
  \item \textbf{Variance and threshold levels:} For our video synchronization, we started and ended each recording with a few seconds in which the recording smartphone was kept still in the air.
  Then we shake the smartphone to get a sudden change in the inertial sensor data, so the video and sensor data stream can easily be aligned.
  During the still period, the data variance, and thus the constructed hypersphere, becomes very small.
  As a results, even small movements are considered to be changes and in the case of a mid-air kept smartphone a lot of false change points are detected.
  Due to the merging effect described above, the final real change point (shaking the smartphone) is not discovered.
  \item \textbf{Incorrect weighting:} In our analyses we have used the data from the accelerometer, magnetic field, and rotation sensors.
  In the segments which embodied movement in a circular manner (such as walking and running around a fountain), the corner was not (or at a different time point) discovered.
  However, when we only used the magnetic field and rotation sensors, the corner was correctly discovered.
  Other transitions, such as from walking to running, are harder to discover without the (linear) accelerometer sensor data.
\end{itemize}

A more elaborate discussion with observations and remarks is provided in \Cref{AppendixA}.

\section{Possible improvements}\label{sec:possible_improvements}
From the above section we can extract a few possible improvements for the algorithm.
Most of the observations and remarks are due to a different required sensitivity for local segments of the run.
Currently, our method employs a global setting for the thresholds.
A possible improvement would be to use locally optimized parameters, although that would require an reflective data processing approach or more a priori information about the data.
The design decision for our approach was to create a general method, which requires no a priori knowledge about the data distribution.

Furthermore, we discovered that sometimes, especially with movements incorporating circular motions, the addition of (linear) accelerometer data makes the discovery of change points harder.
This problem can be handled by, for example, assigning weights to the sensor data streams.
Change can then be detected when it only occurs in one of the streams.

### 2014-01-20 Chapters/Chapter7.tex
Final 1.0

This research has applied \gls{occ} methods to inertial sensor data, recorded during the performance of various human activities, in order to create a temporal segmentation of those recordings.
The goal was to find the change points between activities, assuming that it will aid our understanding of the sensor values.
To do so, we used the \gls{svdd} algorithm by Tax~\cite{tax1999support}, following the method of applying it to a sliding window of data, as done by Camci~\cite{camci2010change}.
From these methods we have created our own method \gls{ocs-hats} and made it publicly available \cite{vlasveld2014hats}, which is a slightly modified implementation of \gls{svcpd} by Camci~\cite{camci2010change} applied to \gls{har} data.
% Instead of using both the information whether a new data point is an outlier and the radius of the hypersphere to indicate change, we only used the latter model properties.
An increase of the radius of the constructed hypersphere, which encloses the data objects and increases with heterogeneous segments of data, is extracted and used for indication of change.

In the following section we will further discuss the proposed method for temporal segmentation of inertial sensor signals from human activities.
Our main findings and contributions are discussed in \Cref{sec:main_findings}.
The limitations of our method and research is subject of \Cref{sec:limitations}.
\Cref{sec:further_research} provides suggestions for future research.
\Cref{sec:final_remarks} is the final section of this thesis.

\section{What is done?}
The goal of this research was to find a temporal segmentation of recordings obtained from human activities.
This was performed in the wider context of activity classification.
Currently, many algorithms obtain an implicit segmentation as a side-effect of direct activity classification.
In this research the primary goal was finding a temporal segmentation, which is assumed to be able to aid the classification step.
This problem is reduced to finding change points in temporal data sets originating from inertial sensors.

@@

In this research we have shown that it is possible to explicitly obtain a temporal segmentation, expressed as change points, of activities performed by humans.
This was performed in a real-world setting, showing the robustness of the method for noisy data.
Since manual labeling (the annotation of change points on the time series data) of human activities is ambiguous we put emphasis on the qualitative inspection of the results.
The proposed method gives fairly good results, although parameter tuning is a subject of further research.

@@

This research is, as far as we know, the first application of \gls{oc-svm} based methods to on-body worn inertial sensor signals, such as accelerometer data.
Furthermore, it applies the algorithms, amongst others of Tax and Duin~\cite{tax1999support}, to real-world data.
We have shown that the assumption of a change the performed activity results, due to a change in the data distribution, in an increase of the hypersphere's radius.
This increase, and a similar decrease in the case of homogeneous activity distribution in the current window, can be used for an indication of change.

The proposed \gls{ocs-hats} method processes the raw inertial sensor data, originating from sensors as accelerometers and magnetic field sensors.
This in contrast with many other approaches that extract features in a pre-processing stage.
The method processes the data in an online manner, which makes it suitable for applications that need direct segmentation of time series data.

Finally, due to encountered problems with commonly used data sets \cite{kwapisz2011activity,anguita2012human}, we have recorded our real-world data sets by ourselves.
We have recorded all the activities with a smartphone worn in the right front pocket of the subjects.
Furthermore, we have recorded the performing subjects from the third person point of view with a video camera.
This allows for better interpretation of the results, since it provides a ground truth which can be consulted in case of ambiguity.
All the recorded data is made publicly available as the \gls{almende-data} data set \cite{vlasveld2014acras}.
Using our results as a benchmark, it is possible for others to use our data set and improve on it.

@@

For the implementation of \gls{ocs-hats} we have created and used a \textsc{Matlab} application.
Because of that, we can not make reliable claims on the speed and memory usage of the algorithm.
With the online nature and applications in mind and different targeted hardware platforms, this is something that could be of importance.

>>>>

This research has applied \gls{occ} methods to inertial sensor data, recorded during the performance of various human activities, in order to create a temporal segmentation of those recordings.
The goal was to find the change points between activities, assuming that it will aid our understanding of the sensor values.
To do so, we used the \gls{svdd} algorithm by Tax~\cite{tax1999support} and applied it to a sliding window of data, as done by Camci~\cite{camci2010change}.
From these methods we have created our own method \gls{ocs-hats} and made it publicly available \cite{vlasveld2014hats}, which is a slightly modified implementation of \gls{svcpd} by Camci~\cite{camci2010change} applied to \gls{har} data.
% Instead of using both the information whether a new data point is an outlier and the radius of the hypersphere to indicate change, we only used the latter model properties.
An increase of the radius of the constructed hypersphere, which encloses the data objects and increases with heterogeneous segments of data, is extracted and used for indication of change.

In \Cref{sec:proposed_method} we will further discuss the proposed method for temporal segmentation of inertial sensor signals from human activities.
Our main findings and contributions are discussed in \Cref{sec:main_findings}.
The limitations of our method and research is subject of \Cref{sec:limitations}.
\Cref{sec:further_research} provides suggestions for future research.
\Cref{sec:final_remarks} is the final section of this thesis.

\section{Proposed method}\label{sec:proposed_method}
The goal of this research was to find a temporal segmentation of recordings obtained from human activities.
This was performed in the wider context of activity classification.
Currently, many algorithms obtain an implicit segmentation as a side-effect of direct activity classification.
In this research the primary goal was finding a temporal segmentation, which is assumed to be able to aid the classification step.
This problem is reduced to finding change points in temporal data sets originating from inertial sensors.

@@

We have shown that it is possible to explicitly obtain a temporal segmentation, expressed as change points, of activities performed by humans.
This was performed in a real-world setting, showing the robustness of the method for noisy data.
Since manual labeling (the annotation of change points on the time series data) of human activities is ambiguous we put emphasis on the qualitative inspection of the results.
The proposed method gives fairly good results, although parameter tuning is a subject of further research.

@@

The proposed method is, as far as we know, the first application of \gls{oc-svm} based methods to on-body worn inertial sensor signals, such as accelerometer data.
Furthermore, it applies the algorithms, amongst others of Tax and Duin~\cite{tax1999support}, to real-world data.
We have shown that the assumption of a change the performed activity results, due to a change in the data distribution, in an increase of the hypersphere's radius.
This increase, and a similar decrease in the case of homogeneous activity distribution in the current window, can be used for an indication of change.

The proposed \gls{ocs-hats} method processes the raw inertial sensor data, originating from sensors as accelerometers and magnetic field sensors.
This contrasts with many other approaches that extract features in a pre-processing stage.

Furthermore, because the it is an online method, it is suitable for applications that need direct segmentation of time series data.

The commonly used data sets \cite{kwapisz2011activity,anguita2012human} are recorded in a non-continuous manner.
The proposed method requires data streams of continuous activities.
Annotating change points on this kind of time series can be ambigious, so we needed to have a ground truth.
Because of these two shortcomings, we have recorded continuous real-world data sets, together with a third persion point of view video stream.
All the recorded data is made publicly available as the \gls{almende-data} data set \cite{vlasveld2014acras}.
Using our results as a benchmark, it is possible for others to use our data set and improve on it.

@@

We have impplemented \gls{ocs-hats} in \textsc{Matlab}.
Because of that, we can not make reliable claims on the speed and memory usage of the algorithm on real-world devices such as smartphones.
With the online nature and applications in mind and different targeted hardware platforms, this is something that could be of importance.

### 2014-01-20 main.tex
Final 1.0

\textsc{\LARGE \univname}\\[1.5cm] % University name
\textsc{\Large Master's Degree Thesis}\\[0.5cm] % Thesis type

\HRule \\[0.4cm] % Horizontal line
{\huge \bfseries \ttitle}\\[0.4cm] % Thesis title
\HRule \\[1.5cm] % Horizontal line

\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
\href{http://rvlasveld.github.io}{\authornames}, \href{http://www.uu.nl}{\textsc{UU}} % Author name - remove the \href bracket to remove the link
\end{flushleft}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{flushright} \large
\emph{Supervisors:} \\
\href{http://www.cs.uu.nl/staff/gv.html}{Dr. G.A.W. \textsc{Vreeswijk}, \href{http://www.uu.nl}\textsc{UU}}  \\ % Supervisor name - remove the \href bracket to remove the link
\href{http://www.annevanrossum.com}{A.C. \textsc{van Rossum}, \href{http://www.almende.com}\textsc{Almende B.V.}} % Supervisor name - remove the \href bracket to remove the link
\end{flushright}
\end{minipage}\\[3cm]

\large \textit{A thesis submitted in fulfilment of the requirements\\ for the degree of \degreename}\\[0.3cm] % University requirement text
\textit{in the}\\[0.4cm]
\facname\\\deptname\\[2cm] % Research group name and department name

{\large \today}\\[4cm] % Date
%\includegraphics{Logo} % University/department logo - uncomment to place it

>>>>

% \begin{figure}[h]
  % \centering
    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/UU-logo2011_PANTONE.eps}
% \end{figure}

\textsc{\LARGE \univname}\\[0.5cm] % University name
\textsc{\Large Master's Degree Thesis}\\[0.5cm] % Thesis type

\HRule \\[0.4cm] % Horizontal line
{\huge \bfseries \ttitle}\\[0.4cm] % Thesis title
\HRule \\[1cm] % Horizontal line

\begin{minipage}[t]{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
\href{http://rvlasveld.github.io}{\authornames}, \href{http://www.uu.nl}{\textsc{UU}} % Author name - remove the \href bracket to remove the link
\end{flushleft}
\end{minipage}
\begin{minipage}[t]{0.5\textwidth}
\begin{flushright} \large
\emph{Supervisors:} \\
\href{http://www.cs.uu.nl/staff/gv.html}{Dr. G.A.W. \textsc{Vreeswijk}, \href{http://www.uu.nl}\textsc{UU}}  \\ % Supervisor name - remove the \href bracket to remove the link
\href{http://www.annevanrossum.com}{A.C. \textsc{van Rossum}, \href{http://www.almende.com}\textsc{Almende B.V.}} % Supervisor name - remove the \href bracket to remove the link
\end{flushright}
\end{minipage}\\[1cm]

\large \textit{A thesis submitted in fulfilment of the requirements\\ for the degree of \degreename}\\[0.3cm] % University requirement text
\textit{in the}\\[0.4cm]
\facname\\\deptname\\[1cm] % Research group name and department name

{\large \today}\\[0.5cm] % Date
\includegraphics{./Figures/faculty_of_science.png} % University/department logo - uncomment to place it
